Diverse Genotype-by-Weather Interactions in Switchgrass

### 2024-06-22 manuscript/revision2/revision2.tex
revision2: Add additional recent GxE analyses as references

Plant phenological timings are major components of plant fitness
affected by multiple external environmental cues (e.g.~degree of winter
chilling, day length, temperature, and water availability) that signal
existing or upcoming growing conditions (1--3). Genetic responses to
environmental cues determine the speed, timing, and energy apportioned
to vegetative and reproductive growth and shape both the individual's
lifespan and its lifetime production of viable seed. Day length (or
photoperiod) is one of the most predictable environmental cues, and
genetic sensitivity to photoperiod protects plants from potentially
fatal consequences of phenological responses to temperature cues at the
``wrong'' time of year. However, the usefulness of specific
environmental cues depends on both features of the environment, such as
cue predictability and relevance, and the species' adaptive strategies
(4). Species with wide natural distributions can have multiple distinct
environmentally cued phenological responses: for example, populations of
sunflower (\emph{Helianthus annuus}) exhibit day-neutral, facultative
short day, and facultative long-day flowering responses, which vary with
their environments (5, 6). Distinct genetic responses in different
environments are known as genotype-by-environment interactions, or GxE.

Flowering time, or the transition from vegetative to reproductive
growth, is a common subject of GxE research (5--11), a key output of
selection driving adaptation to local environments (2, 12, 13), and a
selection target for crop improvement to adapt crops to local or future
environments (14). Changing flowering responsiveness to photoperiod cues
has allowed geographic range expansion and increased yields in several
cereal species (15--19) and other crops (20, 21). Recent statistical
advances in studying phenological GxE have involved determining critical
environmental indices before the phenological event occurs, such as
photothermal time within a critical growth window (10). However, most
studies of flowering GxE focus on finding a single, best fitting form of
genotype-environment covariance, despite the key expectation that
different genetic subpopulations, and even different genomic regions,
have likely evolved distinct patterns of GxE (22). Additionally, despite
theoretical predictions that local adaptation should involve
antagonistic pleiotropy, or alleles with effects with opposing fitness
outcomes (23--26), previous empirical work has found limited evidence of
antagonistic pleiotropy (12, 27). However, this work has been limited by
a known statistical bias that reduced detection of genetic effects that
differ in sign (27--29). Thus, despite substantial interest in the
frequencies of various forms of GxE, the frequency of sign-changing GxE
relative to other forms of GxE remains unknown.

Previous research suggests that switchgrass phenological timings should
have GxWeather and that these timings could differ by genetic
subpopulation. Switchgrass is considered a short-day plant with
reproductive development strongly linked to day of the year (30).
However, as part of its wide environmental adaptation across the eastern
half of North America, its photoperiodicity has been predicted to differ
by plant latitude of origin (31, 32). We previously found divergent
Midwest and Gulf genetic subpopulations of switchgrass with distinct
sets of environmental adaptations, in that both populations had distinct
genetic variation associated with each of two fitness proxies, biomass
and overwinter survival (33). The Midwest genetic subpopulation is
primarily composed of individuals from the well-studied upland
switchgrass ecotype (34, 35), while the Gulf subpopulation has
individuals from the well-studied lowland ecotype and the phenotypically
intermediate coastal ecotype (33).

Here, we test for GxWeather for two phenological timings by assigning
patterns of genetic effects on phenology at eight common gardens to many
patterns of weather covariance at these gardens. To do this, we
phenotype a diversity panel of hundreds of switchgrass genotypes from
the Midwest and Gulf subpopulations for the timing of vegetative and
reproductive development. We do this at eight common garden locations
spanning 17 degrees of latitude: these gardens cover the majority of the
latitudinal and climatic range of switchgrass and capture the most
comprehensive picture to date of the environmental variation this
species encounters. We define multiple ways phenological timings might
covary with weather (Table 1) and additional ways phenological timings
might vary by garden (SI Appendix, Section S1), then jointly re-estimate
genetic effects on these timings at all eight common gardens using the
set of these covariance matrices that significantly improved the modeled
log-likelihood when included (SI Appendix, Section S2-S5) (36). We use
the Bayesian framework \emph{mash} (multivariate adaptive shrinkage)
developed by (36), to refine effect size estimates from genome-wide
association (GWAS). \emph{Mash} allows us to identify and specify
multiple covariance structures among genetic effect estimates across
sites, including structures that represent covariance in weather
variables of interest. Importantly, this method circumvents statistical
biases in detecting genetic effects with the same or opposite signs
(37). To confirm our genetic mapping of GxWeather, we compare genomic
locations of the significant posterior effect estimates from \emph{mash}
to mapping results from an outbred mapping population grown at the same
sites. Our analyses allow us to describe the weather cues and types of
GxE affecting phenology in two divergent natural populations of
switchgrass across the species' latitudinal range.

@@

We expected to observe common rank-changing GxE at the level of
individual loci as this is a key theoretical prediction of local
adaptation (23--26). To determine the frequency of rank-changing GxE, we
used the local false sign rate (lfsr), an analogue of the lfdr that
establishes confidence in the effect sign, not the effect's difference
from zero, to determine significance. We required lfsr significance (p
\textless{} 0.05) in both gardens to include effects. This means that
our tests for a sign change between gardens carry an equal statistical
burden to those for effects with the same sign. We separated kinds of
effects at the level of individual loci into SNP effects that differ in
sign between gardens (effects with rank-changing GxE between gardens),
SNP effects that differ in magnitude (effects that are large in one
garden or region and smaller in others), and SNP effects that are
indistinguishable in two gardens (Figure~\ref{fig-effects} A).

@@

We sought additional experimental support the occurrence and location of
our significant SNP effect re-estimates using an independent pseudo-F2
mapping population created from Gulf \& Midwest individuals and grown at
the same sites (Figure~\ref{fig-qtl} A,B). We conducted quantitative
trait loci (QTL) mapping of flowering as functions of five environmental
cues that we also used as covariance matrices in \emph{mash}, and
identified eight QTL for flowering date, eight QTL for flowering as a
function of day length change two days prior, one QTL for the start of
vegetative growth, and two QTL for vegetative growth as a function of
daylength change one day prior, all of which showed QTL by environment
interactions (SI Appendix, Figure S4). All QTL for flowering overlapped
one or more homologs from rice or \emph{A. thaliana} with functionally
validated roles in flowering (SI Appendix, Dataset 7); the QTL on Chr04K
overlapped a gene recently functionally validated for flowering in
switchgrass (38). All flowering and green-up QTL intervals contained at
least one SNP significant in at least one \emph{mash} run at a
log10-transformed Bayes Factor \textgreater{} 2, or in the 1\% tail of
significance, whichever was stricter (SI Appendix, Dataset 8). We also
looked for enrichments of \emph{mash} SNPs in the 1\% tail of
significance (the `\emph{mash} 1\% tail') within each QTL interval. At
the 5\% level, three QTL had enrichments of SNPs in the \emph{mash} 1\%
tail. Overall, there were five significant enrichments (p \textless{}
0.05, hypergeometric test) of SNPs in the \emph{mash} 1\% tail in the
QTL intervals. Thus, we were able to experimentally support the genomic
windows of some re-estimates of significant SNP effects with a QTL
mapping experiment using a separate mapping population.

@@

Our analysis of the timing of vegetative growth in the Gulf and in Both
subpopulations revealed many alleles with rank-changing GxE, or sign
changes in their effects between the Texas and North gardens
(Figure~\ref{fig-effects} B). As phenological timings are major
components of plant fitness, this result supports theoretical models
that local adaptation should involve antagonistic pleiotropy at the
level of individual loci (23--26). In addition, it expands the genetics
of local adaptation research to more than two field sites and to a wider
range of genetic variation by re-estimating effects from GWAS to
determine that rank-changing GxE for phenological timings is common in
small genomic regions (12, 39, 40).

@@

In 2019, we scored two phenological events every two days in two mapping
populations of switchgrass, a diversity panel and a pseudo-F2 cross,
planted at eight common garden locations (35, 40, 41). We scored the
onset of vegetative growth, or green-up date, as the day of the year
when 50\% of the tiller area of the crown of the plant cut the previous
year had green growth. The onset of reproductive growth, or flowering
date, was the day of the year when 50\% of the plant tillers had
panicles undergoing anthesis.

The formation and resequencing of the diversity panel has been described
previously (33). The diversity panel contained 134 sequenced, clonally
propagated individuals from the Midwest genetic subpopulation, and 229
from the Gulf genetic subpopulation. To allow for the possibility that
different subpopulations had different strengths of connection between
our phenotypes and genotypes (42), we conducted three sets of genetic
analyses: on Gulf and Midwest genotypes separately, and on both
subpopulations together (`Both' subpopulations). Analyses to determine
narrow-sense heritability (h\textsuperscript{2}) for green-up and
flowering were done using linear mixed models and followed (33) (SI
Appendix, Section S6).

To confirm candidate genomic regions found using \emph{mash} on the
diversity panel, we analyzed flowering in an outbred pseudo-F2 cross
between four individuals, two Midwest and two Gulf individuals. The
formation of this mapping population has been described previously (35);
additional details on QTL mapping can be found in SI Appendix, Section
S7. To be directly comparable to the diversity panel data, only 2019
phenology data from the pseudo-F2 cross from the same eight common
garden sites were used.

@@

To use \emph{mash} on our diversity panel, we had to specify both a
relatively uncorrelated set of covariance matrices, which in our case
defined types of GxE and GxWeather between gardens, and we had to
specify subsets of SNP effect estimates and standard errors for our
traits at each common garden. To specify a set of covariance matrices,
we first defined many covariance matrices, including GxWeather matrices
that represented the correlation in weather cues between gardenes before
the phenological event (SI Appendix, Section S1), then implemented a
model selection approach that used a greedy algorithm to evaluate if the
log likelihood of the \emph{mash} model was significantly improved as
additional covariance matrices were included (SI Appendix, Section S4).
To specify subsets of SNP effect estimates and standard errors to use
for both the greedy algorithm and with the optimal set of covariance
matrices, we first calculated best linear unbiased predictors (BLUPs)
for each phenological trait in each genetic subpopulation and each
common garden (SI Appendix, Section S2). Next, we determined effect
estimates for 8.8 to 12.3 million SNPs per subpopulation by conducting
garden-specific GWAS on these BLUPs using \(k\) vectors of singular
values to correct for population structure, where \(k\) was the smallest
integer value that made the genomic control coefficient closest to 1 (SI
Appendix, Section S2). Singular values were computed using singular
value decomposition of the matrix of all SNPs, with iterative SNP
pruning and removal of regions in long-range linkage disequilibrium
(43). Third, to make the \emph{mash} models computationally feasible, we
extracted two subsets of SNP effect estimates and standard errors from
our GWAS effect estimates: (i) effects from a subset of ``strong'' tests
corresponding to stronger effects on our traits; (ii) results from a
random subset of all tests to correspond to an unbiased representation
of all effects (SI Appendix, Section S3). We used the subset of random
effects in our greedy \emph{mash} algorithm (SI Appendix, Section S4)
and used the subset of strong effects in our \emph{mash} models with the
optimal set of covariance matrices (SI Appendix, Section S5).

The loadings of genetic effects onto the multiple covariance structures
specified in our \emph{mash} models provided information on genome-wide
patterns of SNP-environment interaction. In addition, the the GxWeather
covariance structures allowed hypothesis testing of specific weather
variables as cues for the start of vegetative and reproductive growth.
Say that a SNP in the gene Flowering Locus C (FLC), a well-known
flowering time regulator, controls flowering in a photoperiod-dependent
manner. In that case, the joint estimate of effects for that SNP could
have a high mixture proportion, or mass, on a covariance matrix created
using a photoperiod-based environmental cue, such as day length at some
interval prior to flowering. In our data, we would infer that the effect
of that SNP on flowering was caused by a response to the environmental
cue used to construct the GxWeather covariance structure with the
largest mass.

@@

Using the lfsr rather than the local false discovery rate (lfdr) is a
critical change in our ability to detect alleles directly contributing
to rank-changing GxE between environments. The lfdr, like other measures
of FDR, focuses on if we have enough evidence to reject the null
hypothesis that an effect j is 0, or that there is a significant effect.
Previous studies of antagonistic pleiotropy (e.g. (40)) have used the
lfdr or equivalent statistical tests to detect antagonistic pleiotropy.
These tests were conservative, in that they required two non-zero
effects of different signs, while tests for differential sensitivity
required only one non-zero effect. This previous work recognized that
this testing bias could lead to undercounting occurrences of
antagonistic pleiotropy (27, 29), and sought to reduce it by permutation
(28). However, using the lfsr to test for allelic effects that differ in
sign does not undercount these occurences, as this statistic answers a
fundamentally different question. For each effect \(j\), the \(lfsr_j\)
is defined as the probability that we make an error in the sign of
effect \(j\) if we were forced to declare the effect positive or
negative (37). Thus, rather than asking ``Are these two effects
different?'' - as we reasonably expect two effects to be, even if this
difference cannot be measured - the local false sign rate answers a more
meaningful question: Can we be confident in the sign of this effect?

@@

\bibitem[\citeproctext]{ref-botero2015evolutionary}
\CSLLeftMargin{4. }%
\CSLRightInline{C. A. Botero, F. J. Weissing, J. Wright, D. R.
Rubenstein, Evolutionary tipping points in the capacity to adapt to
environmental change. \emph{Proceedings of the National Academy of
Sciences} \textbf{112}, 184--189 (2015).}

\bibitem[\citeproctext]{ref-blackman2013interacting}
\CSLLeftMargin{5. }%
\CSLRightInline{B. K. Blackman, Interacting duplications, fluctuating
selection, and convergence: The complex dynamics of flowering time
evolution during sunflower domestication. \emph{Journal of experimental
botany} \textbf{64}, 421--431 (2013).}

\bibitem[\citeproctext]{ref-henry2014transitions}
\CSLLeftMargin{6. }%
\CSLRightInline{L. P. Henry, R. H. Watson, B. K. Blackman, Transitions
in photoperiodic flowering are common and involve few loci in wild
sunflowers (helianthus; asteraceae). \emph{American Journal of Botany}
\textbf{101}, 1748--1758 (2014).}

\bibitem[\citeproctext]{ref-aagren2017adaptive}
\CSLLeftMargin{7. }%
\CSLRightInline{J. Ågren, C. G. Oakley, S. Lundemo, D. W. Schemske,
Adaptive divergence in flowering time among natural populations of
arabidopsis thaliana: Estimates of selection and QTL mapping.
\emph{Evolution} \textbf{71}, 550--564 (2017).}

\bibitem[\citeproctext]{ref-brachi2010linkage}
\CSLLeftMargin{8. }%
\CSLRightInline{B. Brachi, \emph{et al.}, Linkage and association
mapping of arabidopsis thaliana flowering time in nature. \emph{PLoS
genetics} \textbf{6}, e1000940 (2010).}

\bibitem[\citeproctext]{ref-dittmar2014flowering}
\CSLLeftMargin{9. }%
\CSLRightInline{E. L. Dittmar, C. G. Oakley, J. Ågren, D. W. Schemske,
Flowering time QTL in natural populations of arabidopsis thaliana and
implications for their adaptive value. \emph{Molecular ecology}
\textbf{23}, 4291--4303 (2014).}

\bibitem[\citeproctext]{ref-li2018genomic}
\CSLLeftMargin{10. }%
\CSLRightInline{X. Li, T. Guo, Q. Mu, X. Li, J. Yu, Genomic and
environmental determinants and their interplay underlying phenotypic
plasticity. \emph{Proceedings of the National Academy of Sciences}
\textbf{115}, 6679--6684 (2018).}

\bibitem[\citeproctext]{ref-romero2017study}
\CSLLeftMargin{11. }%
\CSLRightInline{J. A. Romero Navarro, \emph{et al.}, A study of allelic
diversity underlying flowering-time adaptation in maize landraces.
\emph{Nature genetics} \textbf{49}, 476--480 (2017).}

\bibitem[\citeproctext]{ref-wadgymar2017identifying}
\CSLLeftMargin{12. }%
\CSLRightInline{S. M. Wadgymar, \emph{et al.}, Identifying targets and
agents of selection: Innovative methods to evaluate the processes that
contribute to local adaptation. \emph{Methods in Ecology and Evolution}
\textbf{8}, 738--749 (2017).}

\bibitem[\citeproctext]{ref-blumel2015flowering}
\CSLLeftMargin{13. }%
\CSLRightInline{M. Blümel, N. Dally, C. Jung, Flowering time regulation
in crops---what did we learn from arabidopsis? \emph{Current opinion in
biotechnology} \textbf{32}, 121--129 (2015).}

\bibitem[\citeproctext]{ref-jung2009flowering}
\CSLLeftMargin{14. }%
\CSLRightInline{C. Jung, A. E. Müller, Flowering time control and
applications in plant breeding. \emph{Trends in plant science}
\textbf{14}, 563--573 (2009).}

\bibitem[\citeproctext]{ref-turner2005pseudo}
\CSLLeftMargin{15. }%
\CSLRightInline{A. Turner, J. Beales, S. Faure, R. P. Dunford, D. A.
Laurie, The pseudo-response regulator ppd-H1 provides adaptation to
photoperiod in barley. \emph{Science} \textbf{310}, 1031--1034 (2005).}

\bibitem[\citeproctext]{ref-faure2012mutation}
\CSLLeftMargin{16. }%
\CSLRightInline{S. Faure, \emph{et al.}, Mutation at the circadian clock
gene EARLY MATURITY 8 adapts domesticated barley (hordeum vulgare) to
short growing seasons. \emph{Proceedings of the National Academy of
Sciences} \textbf{109}, 8328--8333 (2012).}

\bibitem[\citeproctext]{ref-hung2012zmcct}
\CSLLeftMargin{17. }%
\CSLRightInline{H.-Y. Hung, \emph{et al.}, ZmCCT and the genetic basis
of day-length adaptation underlying the postdomestication spread of
maize. \emph{Proceedings of the National Academy of Sciences}
\textbf{109}, E1913--E1921 (2012).}

\bibitem[\citeproctext]{ref-zakhrabekova2012induced}
\CSLLeftMargin{18. }%
\CSLRightInline{S. Zakhrabekova, \emph{et al.}, Induced mutations in
circadian clock regulator mat-a facilitated short-season adaptation and
range extension in cultivated barley. \emph{Proceedings of the National
Academy of Sciences} \textbf{109}, 4326--4331 (2012).}

>>>>

Plant phenological timings are major components of plant fitness
affected by multiple external environmental cues (e.g.~degree of winter
chilling, day length, temperature, and water availability) that signal
existing or upcoming growing conditions (1--4). Genetic responses to
environmental cues determine the speed, timing, and energy apportioned
to vegetative and reproductive growth and shape both the individual's
lifespan and its lifetime production of viable seed. Day length (or
photoperiod) is one of the most predictable environmental cues, and
genetic sensitivity to photoperiod protects plants from potentially
fatal consequences of phenological responses to temperature cues at the
``wrong'' time of year. However, the usefulness of specific
environmental cues depends on both features of the environment, such as
cue predictability and relevance, and the species' adaptive strategies
(5). Species with wide natural distributions can have multiple distinct
environmentally cued phenological responses: for example, populations of
sunflower (\emph{Helianthus annuus}) exhibit day-neutral, facultative
short day, and facultative long-day flowering responses, which vary with
their environments (6, 7). Distinct genetic responses in different
environments are known as genotype-by-environment interactions, or GxE.

Flowering time, or the transition from vegetative to reproductive
growth, is a common subject of GxE research (6--13), a key output of
selection driving adaptation to local environments (2, 14--16), and a
selection target for crop improvement to adapt crops to local or future
environments (17). Changing flowering responsiveness to photoperiod cues
has allowed geographic range expansion and increased yields in several
cereal species (18--22) and other crops (23, 24). Recent statistical
advances in studying phenological GxE have involved determining critical
environmental indices before the phenological event occurs, such as
photothermal time within a critical growth window (11). However, most
studies of flowering GxE focus on finding a single, best fitting form of
genotype-environment covariance, despite the key expectation that
different genetic subpopulations, and even different genomic regions,
have likely evolved distinct patterns of GxE (25). Additionally, despite
theoretical predictions that local adaptation should involve
antagonistic pleiotropy, or alleles with effects with opposing fitness
outcomes (26--29), previous empirical work has found limited evidence of
antagonistic pleiotropy (14, 30, 31). However, this work has been
limited by a known statistical bias that reduced detection of genetic
effects that differ in sign (30, 32, 33). Thus, despite substantial
interest in the frequencies of various forms of GxE, the frequency of
sign-changing GxE relative to other forms of GxE remains unknown.

Previous research suggests that switchgrass phenological timings should
have GxWeather and that these timings could differ by genetic
subpopulation. Switchgrass is considered a short-day plant with
reproductive development strongly linked to day of the year (34).
However, as part of its wide environmental adaptation across the eastern
half of North America, its photoperiodicity has been predicted to differ
by plant latitude of origin (35, 36). We previously found divergent
Midwest and Gulf genetic subpopulations of switchgrass with distinct
sets of environmental adaptations, in that both populations had distinct
genetic variation associated with each of two fitness proxies, biomass
and overwinter survival (37). The Midwest genetic subpopulation is
primarily composed of individuals from the well-studied upland
switchgrass ecotype (38, 39), while the Gulf subpopulation has
individuals from the well-studied lowland ecotype and the phenotypically
intermediate coastal ecotype (37).

Here, we test for GxWeather for two phenological timings by assigning
patterns of genetic effects on phenology at eight common gardens to many
patterns of weather covariance at these gardens. To do this, we
phenotype a diversity panel of hundreds of switchgrass genotypes from
the Midwest and Gulf subpopulations for the timing of vegetative and
reproductive development. We do this at eight common garden locations
spanning 17 degrees of latitude: these gardens cover the majority of the
latitudinal and climatic range of switchgrass and capture the most
comprehensive picture to date of the environmental variation this
species encounters. We define multiple ways phenological timings might
covary with weather (Table 1) and additional ways phenological timings
might vary by garden (SI Appendix, Section S1), then jointly re-estimate
genetic effects on these timings at all eight common gardens using the
set of these covariance matrices that significantly improved the modeled
log-likelihood when included (SI Appendix, Section S2-S5) (40). We use
the Bayesian framework \emph{mash} (multivariate adaptive shrinkage)
developed by (40), to refine effect size estimates from genome-wide
association (GWAS). \emph{Mash} allows us to identify and specify
multiple covariance structures among genetic effect estimates across
sites, including structures that represent covariance in weather
variables of interest. Importantly, this method circumvents statistical
biases in detecting genetic effects with the same or opposite signs
(41). To confirm our genetic mapping of GxWeather, we compare genomic
locations of the significant posterior effect estimates from \emph{mash}
to mapping results from an outbred mapping population grown at the same
sites. Our analyses allow us to describe the weather cues and types of
GxE affecting phenology in two divergent natural populations of
switchgrass across the species' latitudinal range.

@@

We expected to observe common rank-changing GxE at the level of
individual loci as this is a key theoretical prediction of local
adaptation (26--29). To determine the frequency of rank-changing GxE, we
used the local false sign rate (lfsr), an analogue of the lfdr that
establishes confidence in the effect sign, not the effect's difference
from zero, to determine significance. We required lfsr significance (p
\textless{} 0.05) in both gardens to include effects. This means that
our tests for a sign change between gardens carry an equal statistical
burden to those for effects with the same sign. We separated kinds of
effects at the level of individual loci into SNP effects that differ in
sign between gardens (effects with rank-changing GxE between gardens),
SNP effects that differ in magnitude (effects that are large in one
garden or region and smaller in others), and SNP effects that are
indistinguishable in two gardens (Figure~\ref{fig-effects} A).

@@

We sought additional experimental support the occurrence and location of
our significant SNP effect re-estimates using an independent pseudo-F2
mapping population created from Gulf \& Midwest individuals and grown at
the same sites (Figure~\ref{fig-qtl} A,B). We conducted quantitative
trait loci (QTL) mapping of flowering as functions of five environmental
cues that we also used as covariance matrices in \emph{mash}, and
identified eight QTL for flowering date, eight QTL for flowering as a
function of day length change two days prior, one QTL for the start of
vegetative growth, and two QTL for vegetative growth as a function of
daylength change one day prior, all of which showed QTL by environment
interactions (SI Appendix, Figure S4). All QTL for flowering overlapped
one or more homologs from rice or \emph{A. thaliana} with functionally
validated roles in flowering (SI Appendix, Dataset 7); the QTL on Chr04K
overlapped a gene recently functionally validated for flowering in
switchgrass (42). All flowering and green-up QTL intervals contained at
least one SNP significant in at least one \emph{mash} run at a
log10-transformed Bayes Factor \textgreater{} 2, or in the 1\% tail of
significance, whichever was stricter (SI Appendix, Dataset 8). We also
looked for enrichments of \emph{mash} SNPs in the 1\% tail of
significance (the `\emph{mash} 1\% tail') within each QTL interval. At
the 5\% level, three QTL had enrichments of SNPs in the \emph{mash} 1\%
tail. Overall, there were five significant enrichments (p \textless{}
0.05, hypergeometric test) of SNPs in the \emph{mash} 1\% tail in the
QTL intervals. Thus, we were able to experimentally support the genomic
windows of some re-estimates of significant SNP effects with a QTL
mapping experiment using a separate mapping population.

@@

Our analysis of the timing of vegetative growth in the Gulf and in Both
subpopulations revealed many alleles with rank-changing GxE, or sign
changes in their effects between the Texas and North gardens
(Figure~\ref{fig-effects} B). As phenological timings are major
components of plant fitness, this result supports theoretical models
that local adaptation should involve antagonistic pleiotropy at the
level of individual loci (26--29). In addition, it advances local
adaptation research by using more than two field sites and a wide range
of genetic variation to determine that rank-changing GxE for
phenological timings is common in small genomic regions (14, 43, 44).

@@

In 2019, we scored two phenological events every two days in two mapping
populations of switchgrass, a diversity panel and a pseudo-F2 cross,
planted at eight common garden locations (39, 44, 45). We scored the
onset of vegetative growth, or green-up date, as the day of the year
when 50\% of the tiller area of the crown of the plant cut the previous
year had green growth. The onset of reproductive growth, or flowering
date, was the day of the year when 50\% of the plant tillers had
panicles undergoing anthesis.

The formation and resequencing of the diversity panel has been described
previously (37). The diversity panel contained 134 sequenced, clonally
propagated individuals from the Midwest genetic subpopulation, and 229
from the Gulf genetic subpopulation. To allow for the possibility that
different subpopulations had different strengths of connection between
our phenotypes and genotypes (46), we conducted three sets of genetic
analyses: on Gulf and Midwest genotypes separately, and on both
subpopulations together (`Both' subpopulations). Analyses to determine
narrow-sense heritability (h\textsuperscript{2}) for green-up and
flowering were done using linear mixed models and followed (37) (SI
Appendix, Section S6).

To confirm candidate genomic regions found using \emph{mash} on the
diversity panel, we analyzed flowering in an outbred pseudo-F2 cross
between four individuals, two Midwest and two Gulf individuals. The
formation of this mapping population has been described previously (39);
additional details on QTL mapping can be found in SI Appendix, Section
S7. To be directly comparable to the diversity panel data, only 2019
phenology data from the pseudo-F2 cross from the same eight common
garden sites were used.

@@

To use \emph{mash} on our diversity panel, we had to specify both a
relatively uncorrelated set of covariance matrices, which in our case
defined types of GxE and GxWeather between gardens, and we had to
specify subsets of SNP effect estimates and standard errors for our
traits at each common garden. To specify a set of covariance matrices,
we first defined many covariance matrices, including GxWeather matrices
that represented the correlation in weather cues between gardenes before
the phenological event (SI Appendix, Section S1), then implemented a
model selection approach that used a greedy algorithm to evaluate if the
log likelihood of the \emph{mash} model was significantly improved as
additional covariance matrices were included (SI Appendix, Section S4).
To specify subsets of SNP effect estimates and standard errors to use
for both the greedy algorithm and with the optimal set of covariance
matrices, we first calculated best linear unbiased predictors (BLUPs)
for each phenological trait in each genetic subpopulation and each
common garden (SI Appendix, Section S2). Next, we determined effect
estimates for 8.8 to 12.3 million SNPs per subpopulation by conducting
garden-specific GWAS on these BLUPs using \(k\) vectors of singular
values to correct for population structure, where \(k\) was the smallest
integer value that made the genomic control coefficient closest to 1 (SI
Appendix, Section S2). Singular values were computed using singular
value decomposition of the matrix of all SNPs, with iterative SNP
pruning and removal of regions in long-range linkage disequilibrium
(47). Third, to make the \emph{mash} models computationally feasible, we
extracted two subsets of SNP effect estimates and standard errors from
our GWAS effect estimates: (i) effects from a subset of ``strong'' tests
corresponding to stronger effects on our traits; (ii) results from a
random subset of all tests to correspond to an unbiased representation
of all effects (SI Appendix, Section S3). We used the subset of random
effects in our greedy \emph{mash} algorithm (SI Appendix, Section S4)
and used the subset of strong effects in our \emph{mash} models with the
optimal set of covariance matrices (SI Appendix, Section S5).

The loadings of genetic effects onto the multiple covariance structures
specified in our \emph{mash} models provided information on genome-wide
patterns of SNP-environment interaction. In addition, the the GxWeather
covariance structures allowed hypothesis testing of specific weather
variables as cues for the start of vegetative and reproductive growth.
Say that our diversity panel contains a SNP in the gene CONSTANS (CO), a
well-known flowering time regulator, and that only one of the alleles
affects the promotion of flowering in a photoperiod-dependent manner. In
that case, the joint estimate of effects for that SNP could have a high
mixture proportion, or mass, on a covariance matrix created using a
photoperiod-based environmental cue, such as day length at some interval
prior to flowering. In our data, we would infer that the effect of that
SNP on flowering was cued by the weather variable used to create the
GxWeather covariance structure.

@@

Using the lfsr rather than the local false discovery rate (lfdr) is a
critical change in our ability to detect alleles directly contributing
to rank-changing GxE between environments. The lfdr, like other measures
of FDR, focuses on if we have enough evidence to reject the null
hypothesis that an effect j is 0, or that there is a significant effect.
Previous studies of antagonistic pleiotropy (e.g. (44)) have used the
lfdr or equivalent statistical tests to detect antagonistic pleiotropy.
These tests were conservative, in that they required two non-zero
effects of different signs, while tests for differential sensitivity
required only one non-zero effect. This previous work recognized that
this testing bias could lead to undercounting occurrences of
antagonistic pleiotropy (30, 33), and sought to reduce it by permutation
(32). However, using the lfsr to test for allelic effects that differ in
sign does not undercount these occurences, as this statistic answers a
fundamentally different question. For each effect \(j\), the \(lfsr_j\)
is defined as the probability that we make an error in the sign of
effect \(j\) if we were forced to declare the effect positive or
negative (41). Thus, rather than asking ``Are these two effects
different?'' - as we reasonably expect two effects to be, even if this
difference cannot be measured - the local false sign rate answers a more
meaningful question: Can we be confident in the sign of this effect?

@@

\bibitem[\citeproctext]{ref-maple2024}
\CSLLeftMargin{4. }%
\CSLRightInline{R. Maple, P. Zhu, J. Hepworth, J.-W. Wang, C. Dean,
\href{https://doi.org/10.1093/plphys/kiae109}{Flowering time: From
physiology, through genetics to mechanism}. \emph{Plant Physiology}
\textbf{195}, 190--212 (2024).}

\bibitem[\citeproctext]{ref-botero2015evolutionary}
\CSLLeftMargin{5. }%
\CSLRightInline{C. A. Botero, F. J. Weissing, J. Wright, D. R.
Rubenstein, Evolutionary tipping points in the capacity to adapt to
environmental change. \emph{Proceedings of the National Academy of
Sciences} \textbf{112}, 184--189 (2015).}

\bibitem[\citeproctext]{ref-blackman2013interacting}
\CSLLeftMargin{6. }%
\CSLRightInline{B. K. Blackman, Interacting duplications, fluctuating
selection, and convergence: The complex dynamics of flowering time
evolution during sunflower domestication. \emph{Journal of experimental
botany} \textbf{64}, 421--431 (2013).}

\bibitem[\citeproctext]{ref-henry2014transitions}
\CSLLeftMargin{7. }%
\CSLRightInline{L. P. Henry, R. H. Watson, B. K. Blackman, Transitions
in photoperiodic flowering are common and involve few loci in wild
sunflowers (helianthus; asteraceae). \emph{American Journal of Botany}
\textbf{101}, 1748--1758 (2014).}

\bibitem[\citeproctext]{ref-aagren2017adaptive}
\CSLLeftMargin{8. }%
\CSLRightInline{J. Ågren, C. G. Oakley, S. Lundemo, D. W. Schemske,
Adaptive divergence in flowering time among natural populations of
arabidopsis thaliana: Estimates of selection and QTL mapping.
\emph{Evolution} \textbf{71}, 550--564 (2017).}

\bibitem[\citeproctext]{ref-brachi2010linkage}
\CSLLeftMargin{9. }%
\CSLRightInline{B. Brachi, \emph{et al.}, Linkage and association
mapping of arabidopsis thaliana flowering time in nature. \emph{PLoS
genetics} \textbf{6}, e1000940 (2010).}

\bibitem[\citeproctext]{ref-dittmar2014flowering}
\CSLLeftMargin{10. }%
\CSLRightInline{E. L. Dittmar, C. G. Oakley, J. Ågren, D. W. Schemske,
Flowering time QTL in natural populations of arabidopsis thaliana and
implications for their adaptive value. \emph{Molecular ecology}
\textbf{23}, 4291--4303 (2014).}

\bibitem[\citeproctext]{ref-li2018genomic}
\CSLLeftMargin{11. }%
\CSLRightInline{X. Li, T. Guo, Q. Mu, X. Li, J. Yu, Genomic and
environmental determinants and their interplay underlying phenotypic
plasticity. \emph{Proceedings of the National Academy of Sciences}
\textbf{115}, 6679--6684 (2018).}

\bibitem[\citeproctext]{ref-romero2017study}
\CSLLeftMargin{12. }%
\CSLRightInline{J. A. Romero Navarro, \emph{et al.}, A study of allelic
diversity underlying flowering-time adaptation in maize landraces.
\emph{Nature genetics} \textbf{49}, 476--480 (2017).}

\bibitem[\citeproctext]{ref-bustos-korts2019}
\CSLLeftMargin{13. }%
\CSLRightInline{D. Bustos-Korts, \emph{et al.},
\href{https://doi.org/10.1111/tpj.14414}{Exome sequences and
multi{-}environment field trials elucidate the genetic basis of
adaptation in barley}. \emph{The Plant Journal} \textbf{99}, 1172--1191
(2019).}

\bibitem[\citeproctext]{ref-wadgymar2017identifying}
\CSLLeftMargin{14. }%
\CSLRightInline{S. M. Wadgymar, \emph{et al.}, Identifying targets and
agents of selection: Innovative methods to evaluate the processes that
contribute to local adaptation. \emph{Methods in Ecology and Evolution}
\textbf{8}, 738--749 (2017).}

\bibitem[\citeproctext]{ref-blumel2015flowering}
\CSLLeftMargin{15. }%
\CSLRightInline{M. Blümel, N. Dally, C. Jung, Flowering time regulation
in crops---what did we learn from arabidopsis? \emph{Current opinion in
biotechnology} \textbf{32}, 121--129 (2015).}

\bibitem[\citeproctext]{ref-fournier-level2016}
\CSLLeftMargin{16. }%
\CSLRightInline{A. Fournier-Level, \emph{et al.},
\href{https://doi.org/10.1073/pnas.1517456113}{Predicting the
evolutionary dynamics of seasonal adaptation to novel climates in
{\emph{Arabidopsis thaliana}}}. \emph{Proceedings of the National
Academy of Sciences} \textbf{113} (2016).}

\bibitem[\citeproctext]{ref-jung2009flowering}
\CSLLeftMargin{17. }%
\CSLRightInline{C. Jung, A. E. Müller, Flowering time control and
applications in plant breeding. \emph{Trends in plant science}
\textbf{14}, 563--573 (2009).}

\bibitem[\citeproctext]{ref-turner2005pseudo}
\CSLLeftMargin{18. }%
\CSLRightInline{A. Turner, J. Beales, S. Faure, R. P. Dunford, D. A.
Laurie, The pseudo-response regulator ppd-H1 provides adaptation to
photoperiod in barley. \emph{Science} \textbf{310}, 1031--1034 (2005).}
